An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-25 paper/paper.tex
Paper: reduce TODO count

% TODO make title page look nicer
\begin{titlepage}
\begin{center}
\thispagestyle{empty}
% TODO maybe nice pic? \includegraphics [width = 0.3\textwidth] {images/wpilogo}\\[0.3cm]
% TODO install \textsc font
\Large{\textbf{A Major Qualifying Project Report\\ \large{ON}}}\\[0.7cm]
\LARGE{\textsc {\textbf{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}}}\\[0.5cm]
\vspace{0.2cm}
\Large{\textbf{\\Submitted to the Faculty of }}
\LARGE{\textbf{\\WORCESTER POLYTECHNIC INSTITUTE\\}}
\vspace{0.5cm}
\Large{\textbf{\\In Partial Fulfillment of the Requirement for the}}
\Large{\textbf{\\Degree of Bachelor of Science}}
\vspace{0.5cm}
\Large{\textbf{\\by}}\\[0.5cm]
\large{\textbf{Dan Bouffard}}\\
\large{\textbf{Marc Green}}\\
\vspace{0.5cm}
\large{\textbf{UNDER THE GUIDANCE OF}}\\
\large{\textbf{Professor G\'abor N. S\'ark\"ozy}}\\
\vspace{1cm}
\large{\textbf{\\May 1, 2014}}\\
\end{center}
\end{titlepage}

@@

The number of replicas to maintain will have an impact on the performance of the system. If the replica count is too low, the probability of recovering the  file is lowered. If the replica count is too high, then storage space and bandwidth will be needlessly used. Furthermore, if the system requires a peer to offer a proportional amount of storage space to the amount that it is using on others, the peer will also have to offer a lot more space than what it is using locally for its own data. % TODO this last sentence might go better in the fairness section

@@

% todo reword
This section discusses the infrastructure and the protocols of our system that are concerned with moving data between entities in the network. Henceforth, we use the following terminology to disambiguate participants:

@@

Note the emphasis of the action taken by each: a peer is said to \textit{back up} data, and a node is said to \textit{store} a replica of the peer's data.

PeerStore showed that separating the metadata layer from the data layer can result in significant bandwidth reduction. By relaxing the requirement to strictly maintain a set number of replicas per file, the system can reduce the data migration cost. For example, it is unnecessary to \textit{immediately} find a new replication node when a current replication node goes offline; only when the node has shown to be consistently unavailable should the system seek a replacement. Given that there exists more than one replica of a peer's data, having one temporarily offline is not catastrophic.\footnote{Note that this implies the low possibility of not being able to recover data that is backed up. Also note that this situation is unavoidable in \textbf{all} backup systems. We leverage this fact by relaxing the aforementioned requirement like PeerStore; this sacrifices theoretical availability for performance and usability. See Section \ref{subsec:TheMetadataLayer_sec:SystemDesign} for the discussion of the parameterization of these variables in our implementation. See Section \ref{} for our methodology of testing the parameter. Finally, see Section \ref{} for a comparison of the results of different values for the parameter.}
% TODO add references once we know section names

@@

When a peer determines that a node is consistently unreliable, it does two things. The first is that it adds its own ID to the node's blacklist in the DHT. The second is to move its replica to a new node. The peer sends an update to the DHT that it is removing the replica on the unreliable node. This accomplishes two things: all other peers in the network can see that the node is storing less data, and the node can see that it no longer needs to store the replica, so it can remove it the next time it checks the DHT for any changes. The peer will then look for a new node to put a replica on, using the same process that was used for creating the original backups (see Section \ref{subsec:TheMetadataLayer_sec:SystemDesign}). Since this process uses a randomly generated salt to help pick the ID of the new node, each node has an equal chance of being selected, assuming that it has enough space to store the replica and is not blacklisted by too many peers. This helps to prevent some nodes from being left out of the selection process. % TODO this last sentence can be removed without loss of information, but it might mean that the salt explanation in the crypto section was for nothing.

@@

% TODO: Is it actually called SZTAKI Cloud? I've looked through my emails, and Mihaly never calls it exactly that. If we could put as much as we want, it might be nice to put a screenshot of the web interface.
In order to test our system, we needed an environment that would allow us to create and populate our network with as many nodes as we needed for each test scenario. We used SZTAKI Cloud, a virtualization platform that runs on the enterprise cloud computing system \textit{OpenNebula}. SZTAKI Cloud allowed us to create and clone virtual machines for each test scenario. This is done via the web interface. % TODO: Elaborate more here, depending on how much Mihaly wants us to talk about it.

@@

% TODO what is happening after the 11 minute mark? why are they all uploading again? it might help to look at the download speed graph of everyone

@@

Section \ref{sec:EffectsofChurn} explained that, in the presence of an average amount of churn,

@@

% TODO explain what NAT is why and it makes P2P hard?

@@

\begin{enumerate}
    \item Replace the centralized tracker with a DHT to become completely decentralized.
    \item Encrypted communication between peers to reduce information leakage.
    \item The ability to backup folders, not just files.
\end{enumerate}

>>>>

% TODO make title page look nicer
\begin{titlepage}
\begin{center}
\thispagestyle{empty}
% TODO install \textsc font
\Large{\textbf{A Major Qualifying Project Report\\ \large{ON}}}\\[0.7cm]
\LARGE{\textsc {\textbf{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}}}\\[0.5cm]
\vspace{0.2cm}
\Large{\textbf{\\Submitted to the Faculty of }}
\LARGE{\textbf{\\WORCESTER POLYTECHNIC INSTITUTE\\}}
\vspace{0.5cm}
\Large{\textbf{\\In Partial Fulfillment of the Requirement for the}}
\Large{\textbf{\\Degree of Bachelor of Science}}
\vspace{0.5cm}
\Large{\textbf{\\by}}\\[0.5cm]
\large{\textbf{Dan Bouffard}}\\
\large{\textbf{Marc Green}}\\
\vspace{0.5cm}
\large{\textbf{UNDER THE GUIDANCE OF}}\\
\large{\textbf{Professor G\'abor N. S\'ark\"ozy}}\\
\vspace{1cm}
\large{\textbf{\\May 1, 2014}}\\
\end{center}
\end{titlepage}

@@

The number of replicas to maintain will have an impact on the performance of the system. If the replica count is too low, the probability of recovering the file is lowered. If the replica count is too high, then bandwidth and storage space will be needlessly used; the peers storing replicas do not significantly contribute to the availability of the data, and the peer backing up the data will need to store more data for others as part of the fairness mechanism.

@@

This section discusses the data structures and protocols of our system that are used to move data between entities in the network. Henceforth, we use the following terminology to disambiguate participants:

@@

Note the emphasis of the action taken by each: a peer is said to \textit{back up} data, and a node is said to \textit{store} a replica of the peer's data. Understanding this distinction will aid in understanding the description of our system.

PeerStore showed that separating the metadata layer from the data layer can result in significant bandwidth reduction. By relaxing the requirement to strictly maintain a set number of replicas per file, the system can reduce the data migration cost. For example, it is unnecessary to \textit{immediately} find a new replication node when a current replication node goes offline; only when the node has shown to be consistently unavailable should the system seek a replacement. Given that there exists more than one replica of a peer's data, having one temporarily offline is not catastrophic.\footnote{Note that this implies the low possibility of not being able to recover data that is backed up. Also note that this situation is unavoidable in \textbf{all} backup systems. We leverage this fact by relaxing the aforementioned requirement like PeerStore; this sacrifices theoretical availability for performance and usability.}

@@

When a peer determines that a node is consistently unreliable, it does two things. The first is that it adds its own ID to the node's blacklist in the DHT. The second is to move its replica to a new node. The peer sends an update to the DHT that it is removing the replica on the unreliable node. This accomplishes two things: all other peers in the network can see that the node is storing less data, and the node can see that it no longer needs to store the replica, so it can remove it the next time it checks the DHT for any changes. The peer will then look for a new node to put a replica on, using the same process that was used for creating the original backups (see Section \ref{subsec:TheMetadataLayer_sec:SystemDesign}). %Since this process uses a randomly generated salt to help pick the ID of the new node, each node has an equal chance of being selected, assuming that it has enough space to store the replica and is not blacklisted by too many peers. This helps to prevent some nodes from being left out of the selection process. % TODO this last sentence can be removed without loss of information, but it might mean that the salt explanation in the crypto section was for nothing.

@@

% TODO: Is it actually called SZTAKI Cloud? I've looked through my emails, and Mihaly never calls it exactly that.
In order to test our system, we needed an environment that would allow us to create and populate our network with as many nodes as we needed for each test scenario. We used SZTAKI Cloud, a virtualization platform that runs on the enterprise cloud computing system \textit{OpenNebula}. SZTAKI Cloud allowed us to create and clone virtual machines for each test scenario. This is done via the web interface. % TODO: Elaborate more here, depending on how much Mihaly wants us to talk about it.

@@

% TODO what is happening after the 11 minute mark? why are they all uploading again? it might help to look at the download speed graph of everyone
% Dan thinks it just load distribution. explain this

@@

Section \ref{sec:EffectsofChurn} explained that, in the presence of an average amount of churn, the average peer would need to make a new replica of any given file every 11 days.

@@

% TODO briefly explain what NAT is why and it makes P2P hard

@@

\begin{itemize}
    \item Replace the centralized tracker with a DHT to become completely decentralized.
    \item Encrypted communication between peers to reduce information leakage.
    \item The ability to backup folders, not just files.
\end{itemize}
